---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data
BRFSS is a national telephone survey implemented by the CDC. Data is collected from > 400,000 American adults annually.
While it is an exceptionally large sample, its findings cannot be considered universally generalizable for two reasons.

1. Data is collected by phone. Individuals who do not have access to a phone cannot be included.
2. Data is only collected from American adults.

* * *

## Part 2: Research questions

**Research quesion 1:**
Does fruit and vegetable consumption differ reliably by state?

**Research quesion 2:**
Does income predict poor physical or mental health?

**Research quesion 3:**
Does partisan legislative control predict health outcomes for their constituents?

Data sourced from https://ballotpedia.org/Partisan_composition_of_state_legislatures

* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r}
exploreNutrition <- data.frame(brfss2013$X_state, brfss2013$X_frutsum, brfss2013$X_vegesum)
exploreNutrition <- exploreNutrition[complete.cases(exploreNutrition), ]
colnames(exploreNutrition) <- c('state','fruitSum','vegSum')

exploreNutrition <- exploreNutrition %>%
        mutate(nutrientIntake = fruitSum + vegSum)

# Low to High
exploreNutrition %>%
        group_by(state) %>%
        summarise(averageNutrientIntake = mean(nutrientIntake), n = n()) %>%
        arrange(averageNutrientIntake)
```
```{r}
# High to Low
exploreNutrition %>%
        group_by(state) %>%
        summarise(averageNutrientIntake = mean(nutrientIntake), n = n()) %>%
        arrange(desc(averageNutrientIntake))
```

```{r}
ggplot(data = exploreNutrition, aes(x = factor(state), y = nutrientIntake)) + 
        geom_boxplot() + 
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


**Research quesion 2:**

```{r}
exploreIncome <- data.frame(brfss2013$X_state, brfss2013$employ1, brfss2013$income2, brfss2013$poorhlth)
colnames(exploreIncome) <- c('state','employmentStatus','incomeLevel','poorHealth')

exploreIncome <- exploreIncome[complete.cases(exploreIncome), ]

exploreIncome %>%
        group_by(incomeLevel) %>%
        summarise(averagePoorHealthDays = mean(poorHealth)) %>%
        arrange(desc(averagePoorHealthDays))
```

```{r}
ggplot(exploreIncome, aes(x = incomeLevel, y = poorHealth)) +
        geom_boxplot() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 7))
```
These data clearly show a trend in economic status predicting mental and physical health outcomes, such that more income yields lower rates of health problems.

**Research quesion 3:**

```{r}
blueStates <- c('Washington','Oregon','California','Nevada','Colorado','New Mexico',
                'Illinois','Maryland','Delaware','New Jersey','New York','Connecticut',
                'Massachusets','Rhoad Island','Vermont','New Hampshire','Maine')

explorePolitics <- data.frame(brfss2013$X_state, brfss2013$hlthcvrg, brfss2013$lstcovrg, brfss2013$medcost, brfss2013$carercvd)
colnames(explorePolitics) <- c('state','healthCare','timeSinceHealthCare','medCost','careReceived')
explorePolitics <- explorePolitics[explorePolitics$state != c('Puerto Rico','Guam'), ]
explorePolitics <- explorePolitics[complete.cases(explorePolitics), ]

explorePolitics <- explorePolitics %>%
        mutate(statePolitics = ifelse(test = (explorePolitics$state %in% blueStates), yes = 'Democrat', no = 'Republican'))

explorePolitics %>%
        group_by(statePolitics) %>%
        summarise(satisfaction = sum(careReceived == 'Very satisfied'), satisfactionRate = (satisfaction / n()), n = n()) %>%
        arrange(desc(satisfaction))

```

```{r}
ggplot(explorePolitics, aes(x = careReceived, fill = statePolitics)) + 
        geom_bar()

```
These data suggest that Republican states have higher rates of healthcare satisfaction relative to Democratic states.
